---
title: "FITgraphics Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FITgraphics Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

**FITgraphics** is a package designed to take a file created by a fitness tracker or cycling computer,
and allow the user to create aesthetically pleasing graphical representations of their
activity data. 

## Installing package and external dependancies

**FITgraphics** is currently available for download from [GitHub](https://github.com/jstanosheck/FITgraphics), it can be downloaded using
the **devtools** package. 

This package is also dependent on an external package not available on CRAN 
**FITfileR**, so it must also be downloaded using the **devtools** package. 

If the **devtools** package is not yet installed on your machine, this will 
automatically install it.

```{r, message=FALSE, cache=TRUE}
if(!requireNamespace("devtools")) {
    install.packages("devtools")
}

devtools::install_github("grimbough/FITfileR")

devtools::install_github("jstanosheck/FITgraphics")
```

After having installed the needed packages, you should load the libraries to the 
R session to be used.

```{r setup}
library(FITgraphics)
```

# Examples using provided .FIT files 

There are two example data files that come with the **FITgraphics** package, on 
called TestRun.fit and the other called TestRide.fit. These two files can be 
found in the `Data/` directory.

## Loading the two test data files

When loading the files in, you can use the `getFit()` function. It is 
important to note that the files that are loaded into this function must have 
the .fit extension. 

Here we will use both the TestRun.fit and TestRide.fit files to get their location.

```{r, cache=FALSE}
file_location_run <- system.file("extdata", "TestRun.fit", 
                             package = "FITgraphics")

file_location_ride <- system.file("extdata", "TestRide.fit",
                                  package = "FITgraphics")
```

Using the `getFit()` function, you can save each of the two files in 
different variables. Here I will use `runFile` and `rideFile` as 
examples.

```{r, cache=TRUE, message=FALSE}
runFile <- getFit(file_location_run)

rideFile <- getFit(file_location_ride)
```

Each of the two generated variables `runFile` and `rideFile` are a list of three 
items. I will use only `runFile` as an example.

```{r}
names(runFile)
```

Each of these three items are used in the package and can be accessed by the user.


* `"Data"`: a `data.frame` containing all of the data stored in the file 
that is used to form the graphics.
* `"records"`: a `vector` of each of the headings in the `data.frame`.
* `"session"`: a `list` of different metadata about the activity, e.g. total 
elapsed time of the activity or the maximum elevation.


Individual items of each item can be accessed.

```{r}
runFile$session$total_elapsed_time

runFile$session$total_calories

runFile$records
```

## Plotting the files

The `plotFit()` function is used to graph any of the `"records"` against time 
for the activity. A simple example of this is to plot the speed vs time.

`plotFit()` requires two arguments:

* `fitFile` - file that is the output form `getFit()`
* `varName` - string of one of the `"records"`

```{r, cache=TRUE, fig.align='center', fig.height=4, fig.width=6}
plotFit(runFile, "speed")
```

The plot axis, title, units, and graph color will all change based on the 
`varName` argument. The currently supported values of `varName` are:

* `"speed"`
* `"altitude"`
* `"power"`
* `"heart_rate"`
* `"temperature"`
* `"cadence "`

Any other inputs for `varName` that are in the `"records"` will all be plotted 
with the same plot color and an "Unknown" units.

As an example of a different plotting parameters, we can use the `rideFile` to 
plot the power. 

`plotFit()` also can take two optional arguments: 

* `showAverage`
* `showMax`

These arguments are logical arguments with a default of `FALSE`. 


```{r, cache=TRUE, fig.align='center', fig.height=4, fig.width=6}
plotFit(rideFile, "power", showAverage = T, showMax = T)
```

These arguments may cover some of the graph, so it is advisable to graph the 
plot both with and without these arguments if this is an issue.

## Plot the map for each plot

The `mapFit()` function creates a map showing the path of the activity overlayed 
on an open source street map. The starting point of the activity is marked by a 
green marker and the ending point is marked by a red marker. 

```{r, fig.align='center', fig.height=5, fig.width=7}
mapFit(runFile)
```


## Plot the training effect 

The training effect is a value between 0 and 5. Generally, a higher value for 
training effect represents a greater increase in fitness. This metric is developed 
by **Firstbeat** Technologies Ltd. 

The ranges for training effect and how it effects your fitness:

* 0.0 - 0.9: No Benefit 
* 1.0 - 1.9: Minor Benefit
* 2.0 - 2.9: Maintaining
* 3.0 - 3.9: Improving
* 4.0 - 4.9: Highly Improving
* 5.0: Overreaching

There are two types of training effect that can be represented: aerobic and 
anaerobic. Aerobic training effect is generally representative of your ability 
to exercise at a moderate to moderately high intensity for extended periods of time.
Anaerobic capacity is generally representative of your ability to perform high 
intensity activities for a short period of time. 

To plot the training effect we can use `showTrainingEffect()` which takes one 
required argument and one optional argument.

* `fitFile` - (Required) file that is the output form `getFit()`
* `AnaerobicTE` - (Optional) Shows the Anaerobic training effect rather than 
the aerobic training effect


```{r}
#This shows the aerobic training effect for runFile
showTrainingEffect(runFile)
```


```{r, fig.align='center'}
#This shows the anaerobic training effect for runFile
showTrainingEffect(runFile, AnaerobicTE = TRUE)
```
