---
title: "FITgraphics Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FITgraphics Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

**FITgraphics** is a package designed to take a file created by a fitness tracker or cycling computer,
and allow the user to create aesthetically pleasing graphical representations of their
activity data. 

## Installing package and external dependancies

**FITgraphics** is currently available for download from [GitHub](https://github.com/jstanosheck/FITgraphics), it can be downloaded using
the **devtools** package. 

This package is also dependent on an external package not available on CRAN 
**FITfileR**, so it must also be downloaded using the **devtools** package. 

If the **devtools** package is not yet installed on your machine, this will 
automatically install it.

```{r, message=FALSE, cache=TRUE}
if(!requireNamespace("devtools")) {
    install.packages("devtools")
}

devtools::install_github("grimbough/FITfileR")

devtools::install_github("jstanosheck/FITgraphics")
```

After having installed the needed packages, you should load the libraries to the 
R session to be used.

```{r setup}
library(FITgraphics)
```

## Example .FIT files 

There are two example data files that come with the **FITgraphics** package, on 
called TestRun.fit and the other called TestRide.fit. These two files can be 
found in the \code{Data/} directory.

## Loading the two test data files

When loading the files in, you can use the \code{getFit()} function. It is 
important to note that the files that are loaded into this function must have 
the .fit extension. 

```{r, cache=TRUE,  message=FALSE}
file_location_run <- system.file("Data", "TestRun.fit", 
                             package = "FITgraphics")

runFile <- getFit(file_location_run)

file_location_ride <- system.file("Data", "TestRide.fit",
                                  package = "FITgraphics")

rideFile <- getFit(file_location_ride)
```





